### 2015-04-30 article-00-headers.tex
First version submitted

\documentclass[envcountsame]{llncs}

\usepackage{makeidx}    % allows indexgeneration
\usepackage{graphicx}   % allows \includegraphics
%\usepackage[sorting=none]{biblatex}
\usepackage{cite}       % enables autonumbering in cites
\usepackage{titlesec}   % enables \titlespacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Additional libraries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{color}
\usepackage{linegoal}

%
% LISTINGS
%

\usepackage{listings} % see manual: ftp://ftp.funet.fi/pub/TeX/CTAN/macros/latex/contrib/listings/listings.pdf

% language Ruby is used to enable comment lines started with hash character

\lstset{language=Ruby,
aboveskip=6pt,
belowskip=6pt,
backgroundcolor=\color[gray]{0.9},
numberstyle=\scriptsize,
basicstyle=\ttfamily\scriptsize,
numbers=none,
stepnumber=1,
frame=none,
breaklines=true,
breakautoindent=true}

\AtBeginDocument{%
  \renewcommand{\thelstlisting}{\arabic{lstlisting}}
 }

%\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}

% \AtBeginDocument{
%     \newtheoremstyle{principleStyle}
%         {1.5\topsep} % Space above, default \topsep
%         {0.5\topsep} % Space below, default \topsep
%         {\itshape} % Body font
%         {} % Indent amount, example: \parindent
%         {\bfseries\itshape} % Theorem head font
%         {.} % Punctuation after theorem head
%         {.5em} % Space after theorem head
%         {} % Theorem head spec (can be left empty, meaning `normal')

%     \theoremstyle{principleStyle} \newtheorem{principle}{Principle}
% }

\titlespacing*{\section}
    {0pt}{\baselineskip}{0.5\baselineskip}    % {<left>}{<before-sep>}{<after-sep>}

\titlespacing*{\subsection}
    {0pt}{0.8\baselineskip}{0.5\baselineskip}

\titlespacing*{\subsubsection}
    {0pt}{0.8\baselineskip}{0.5em}

%\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}

%
% THEOREMS
%
%\usepackage{amsthm} % enables \newtheorem{}{}

%
% TABLES
%

\usepackage{tabularx} % see manual: http://en.wikibooks.org/wiki/LaTeX/Tables#The_tabularx_package

%\usepackage{multirow}

%\usepackage{svg}        % allows adding scalable vector graphic files

%\usepackage{calc}

\graphicspath{{images/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% User-defined commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \newcommand{\definition}[1]{\textbf{#1}}
%\newcommand{\name}[1]{\emph{#1}}

% programming language names
\newcommand{\name}[1]{\textsf{\footnotesize{}#1}}

% highlight texts
\newcommand{\highlight}[1]{\colorbox{yellow}{\parbox[t]{\linegoal}{#1}}}

% TODO text
% \newcommand{\TODO}[1]{\highlight{\textbf{TODO: }{#1}}}
\newcommand{\TODO}[1]{{}}

%\newcommand{\REVIEW}[1]{\textit{#1}}

% ifcOWL layer names
\newcommand{\bname}[1]{\textnormal{\textbf{#1}}}

\newcommand{\simple}{\bname{Simple}}
\newcommand{\standard}{\bname{Standard}}
\newcommand{\extended}{\bname{Extended}}

\newcommand{\ifcowl}{\bname{ifc\-OWL}}
\newcommand{\ifcrdf}{\bname{ifc\-RDF}}
\newcommand{\ifcsimple}{\ifcowl{}-\simple{}}
\newcommand{\ifcstandard}{\ifcowl{}-\standard{}}
\newcommand{\ifcextended}{\ifcowl{}-\extended{}}

\newcommand{\ifcconverter}{\bname{IFC2LD}}
\newcommand{\ifcld}{\ifcconverter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads
\addtocmark{IFC-EXPRESS-to-OWL-2} % additional mark in the TOC
%

%\vspace{1cm}
%
\title{Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology}
%
\titlerunning{Building ontology}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Nam Vu Hoang \and Seppo T\"orm\"a}
%
%\authorrunning{} % abbreviated author list (for running head)
%
\institute{Department of Computer Science, \\
Aalto University School of Science, 
Espoo, Finland\\
\email{\{nam.vuhoang,seppo.torma\}@aalto.fi}}

\maketitle              % typeset the title of the contribution

\begin{abstract}

The volume of data produced in the construction industry is rapidly growing but, as of now, it is not available for open and flexible access and utilization. We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology. Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components. It paves the way to a distributed and loosely coupled architecture for building information management, and provides opportunities for innovative applications.

\keywords{Ontology, Linked Data, OWL, RDF, IFC, BIM}

\end{abstract}

\input{article-01-introduction.tex}

\input{article-02-main.tex}

\section*{Acknowledgements}

This work has been carried out at Aalto University in research projects DRUM (RYM/PRE 2010-2014) and DRUMBEAT (2014-2017), funded by Tekes, Aalto University, and the participating companies.

\input{article-03-bibliography.tex}

\clearpage
\addtocmark[2]{Author Index} % additional numbered TOC entry
\renewcommand{\indexname}{Author Index}
\printindex
\clearpage
\addtocmark[2]{Subject Index} % additional numbered TOC entry
\markboth{Subject Index}{Subject Index}
\renewcommand{\indexname}{Subject Index}
% \input{subjidx.ind}
\end{document}