### 2015-05-01 article-00-headers.tex
Version 1.2

\documentclass[envcountsame]{llncs}

\usepackage{makeidx}    % allows indexgeneration
\usepackage{graphicx}   % allows \includegraphics

\usepackage{cite}       % enables autonumbering in cites

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Additional libraries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{color}
\usepackage{linegoal}

%
% LISTINGS
%

\usepackage{listings} % see manual: ftp://ftp.funet.fi/pub/TeX/CTAN/macros/latex/contrib/listings/listings.pdf

% language Ruby is used to enable comment lines started with hash character

\lstset{language=Ruby,
aboveskip=6pt,
belowskip=6pt,
backgroundcolor=\color[gray]{0.9},
numberstyle=\scriptsize,
basicstyle=\ttfamily\scriptsize,
numbers=none,
stepnumber=1,
frame=none,
breaklines=true,
breakautoindent=true}

\AtBeginDocument{%
  \renewcommand{\thelstlisting}{\arabic{lstlisting}}
 }

%\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}

% \AtBeginDocument{
%     \newtheoremstyle{principleStyle}
%         {1.5\topsep} % Space above, default \topsep
%         {0.5\topsep} % Space below, default \topsep
%         {\itshape} % Body font
%         {} % Indent amount, example: \parindent
%         {\bfseries\itshape} % Theorem head font
%         {.} % Punctuation after theorem head
%         {.5em} % Space after theorem head
%         {} % Theorem head spec (can be left empty, meaning `normal')

%     \theoremstyle{principleStyle} \newtheorem{principle}{Principle}
% }

\usepackage{titlesec}   % enables \titlespacing

\titlespacing*{\section}
    {0pt}{\baselineskip}{0.5\baselineskip}    % {<left>}{<before-sep>}{<after-sep>}

\titlespacing*{\subsection}
    {0pt}{0.8\baselineskip}{0.5\baselineskip}

\titlespacing*{\subsubsection}
    {0pt}{0.8\baselineskip}{0.5em}

%\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}

%
% THEOREMS
%
%\usepackage{amsthm} % enables \newtheorem{}{}

%
% TABLES
%

\usepackage{tabularx} % see manual: http://en.wikibooks.org/wiki/LaTeX/Tables#The_tabularx_package

%\usepackage{multirow}

%\usepackage{svg}        % allows adding scalable vector graphic files

%\usepackage{calc}

% \graphicspath{{images/}}

\usepackage{caption}    % enables \captionsetup for tables and figures

\AtBeginDocument{
    \captionsetup{font=small}
    \captionsetup{labelfont=bf}
    % \captionsetup[table]{aboveskip=0pt}
    \captionsetup[table]{belowskip=0pt}

    % \setlength{\textfloatsep}{10pt}  % distance between floats on the top or the bottom and the text;
    % \setlength{\floatsep}{0pt}      % distance between two floats;
    \setlength{\intextsep}{\baselineskip}     % distance between floats inserted inside the page text (using h) and the text proper.

%     The default values in the article document class with the 10pt option are:
% \textfloatsep: 20.0pt plus 2.0pt minus 4.0pt;
% \floatsep: 12.0pt plus 2.0pt minus 2.0pt;
% \intextsep: 12.0pt plus 2.0pt minus 2.0pt.
}

%
% REFERENCES
%

% \usepackage[sorting=none]{biblatex} % http://www.ctan.org/tex-archive/help/Catalogue/entries/biblatex.html

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% User-defined commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \newcommand{\definition}[1]{\textbf{#1}}
%\newcommand{\name}[1]{\emph{#1}}

% programming language names
\newcommand{\name}[1]{\textsf{\footnotesize{}#1}}

% highlight texts
\newcommand{\highlight}[1]{\colorbox{yellow}{\parbox[t]{\linegoal}{#1}}}

% TODO text
% \newcommand{\TODO}[1]{\highlight{\textbf{TODO: }{#1}}}
\newcommand{\TODO}[1]{{}}

%\newcommand{\REVIEW}[1]{\textit{#1}}

% ifcOWL layer names
\newcommand{\bname}[1]{\textnormal{\textbf{#1}}}

\newcommand{\simple}{\bname{Simple}}
\newcommand{\standard}{\bname{Standard}}
\newcommand{\extended}{\bname{Extended}}

\newcommand{\ifcowl}{\bname{ifc\-OWL}}
\newcommand{\ifcrdf}{\bname{ifc\-RDF}}
\newcommand{\ifcsimple}{\ifcowl{}-\simple{}}
\newcommand{\ifcstandard}{\ifcowl{}-\standard{}}
\newcommand{\ifcextended}{\ifcowl{}-\extended{}}

\newcommand{\ifcconverter}{\bname{IFC2LD}}
\newcommand{\ifcld}{\ifcconverter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads
\addtocmark{IFC-EXPRESS-to-OWL-2} % additional mark in the TOC
%

%\vspace{1cm}
%
\title{Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology}
%
% \titlerunning{Building ontology}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Nam Vu Hoang \and Seppo T\"orm\"a}
%
%\authorrunning{} % abbreviated author list (for running head)
%
\institute{Department of Computer Science, \\
Aalto University School of Science, 
Espoo, Finland\\
\email{\{nam.vuhoang,seppo.torma\}@aalto.fi}}

\maketitle              % typeset the title of the contribution

\begin{abstract}

The volume of data produced in the construction industry is rapidly growing but, as of now, it is not available for open, flexible access and utilization. We propose a new approach to the automatic derivation of a building ontology from the conceptualization of Industry Foundation Classes (IFC), the existing standard for representing building designs. The resulted multilayer ontology and the tool for converting building designs from IFC files to RDF datasets that conform to the ontology are introduced. Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets using Linked Data tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components. It paves the way to a distributed, loosely coupled architecture for building information management, and provides opportunities for innovative applications.

\keywords{Ontology, Linked Data, OWL, RDF, IFC, BIM}

\end{abstract}

\input{article-01-introduction.tex}

\input{article-02-main.tex}

% According to LNCS manual, the Acknowledgements section must be declared by using \subsubsection{}

\subsubsection{Acknowledgements.} This work has been carried out at Aalto University in research projects DRUM (RYM/PRE 2010-2014) and DRUMBEAT (2014-2017), funded by Tekes, Aalto University, and the participating companies.

%\input{article-03-bibliography.tex}

% \bibliographystyle{splncs03}
\bibliographystyle{splncs}
\bibliography{article.bib}

% \printbibliography{}

\clearpage
\addtocmark[2]{Author Index} % additional numbered TOC entry
\renewcommand{\indexname}{Author Index}
\printindex
\clearpage
\addtocmark[2]{Subject Index} % additional numbered TOC entry
\markboth{Subject Index}{Subject Index}
\renewcommand{\indexname}{Subject Index}
% \input{subjidx.ind}
\end{document}